---
title: "GO_terms"
author:
- name: Nikita Verheyden
  affiliation: JLU Giessen
  date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
      toc: TRUE
      toc_float: TRUE
      code_folding: hide

header-includes:
- \makeatletter\renewcommand*{\fps@figure}{h}\makeatother
- \usepackage{placeins}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Setup

directory
```{r dir}
setwd("C:/Users/nikit/Documents/Krueger_Lab/Publications/miR181_paper/Figure2/06_GO_analysis")
```


## Packages
```{r packages}
source("C:/Users/nikit/Documents/Krueger_Lab/Publications/miR181_paper/Supporting_scripts/themes/theme_paper.R")
library(ggplot2)
library(rtracklayer)
library(dplyr)
library(tibble)
library(Rtsne)
library(cliProfiler)
library(GenomicRanges)
library(readxl)
library(writexl)
library(clusterProfiler)
library(biomaRt)
library(org.Mm.eg.db)
library(msigdbr)
library(eulerr)
```

## Import data

```{r data}
#Ribo profiling
RNA <- read.csv("C:/Users/nikit/Documents/Krueger_Lab/Publications/miR181_paper/Figure3/01_Ribosome_profiling_pipeline/RNA_masterframe.csv")
RPF <- read.csv("C:/Users/nikit/Documents/Krueger_Lab/Publications/miR181_paper/Figure3/01_Ribosome_profiling_pipeline/RPF_masterframe.csv")

#targets
larget <- readRDS("C:/Users/nikit/Documents/Krueger_Lab/Publications/miR181_paper/Figure2/02_compare_miR_enriched_to_Ago_binding_sites/mir181_bs.rds")
largetframe <- as.data.frame(larget)
table(largetframe$set)
largetframe <- largetframe[!(largetframe$set == "ago_bs_mir181_chi"),]
```

## colours
```{r colours}
#colours
farbeneg <- "#b4b4b4"
farbe1 <- "#0073C2FF"
farbe2 <- "#EFC000FF"
farbe3 <- "#CD534CFF"
farbe4 <- "#7AA6DCFF"
farbe5 <- "#868686FF"
farbe6 <- "#003C67FF"
farbe7 <- "#8F7700FF"
farbe8 <- "#3B3B3BFF"
farbe9 <- "#A73030FF"
farbe10 <- "#4A6990FF"
farbe11 <- "#FF6F00FF"
farbe12 <- "#C71000FF"
farbe13 <- "#008EA0FF"
farbe14 <- "#8A4198FF"
farbe15 <- "#5A9599FF"
farbe16 <- "#FF6348FF"
  
RNApcol <- "#b56504"
RNAncol <- "#027d73"
RPFpcol <- "#c4c404"
RPFncol <- "#8d0391"
```



# GO analysis

## enrichGO
-add entrezIDs
-enrichGO analysis
-dotplots for Biological Process (BP), Molecular Function (MF) and Cell Compartmant (CC)
```{r goplots}

#entrezID for binding sites
mart <- useMart("ensembl", dataset = "mmusculus_gene_ensembl", host = "https://jul2019.archive.ensembl.org")
genes <- largetframe$geneID
G_list <- getBM(filters= "ensembl_gene_id", attributes= c("ensembl_gene_id","entrezgene_id"),values=genes,mart= mart)
colnames(G_list) <- c("geneID", "entrezgene_id")
largetframe <- left_join(largetframe,G_list,by="geneID")
head(largetframe)

#entrezID for background (RNAseq)
genesBG <- RNA$Gene
G_listBG <- getBM(filters= "ensembl_gene_id_version", attributes= c("ensembl_gene_id_version","entrezgene_id"),values=genesBG,mart= mart)
colnames(G_listBG) <- c("Gene", "entrezgene_id")
RNA <- left_join(RNA,G_listBG,by="Gene")
head(RNA)


bg <- as.character(RNA[!is.na(RNA$entrezgene_id) & RNA$baseMean > 0,"entrezgene_id"])
l <- as.character(largetframe[!duplicated(largetframe$entrezgene_id) & largetframe$entrezgene_id %in% bg,"entrezgene_id"])


delargBP <- enrichGO(l, OrgDb = org.Mm.eg.db, ont = "BP", universe = bg, pvalueCutoff=0.05, pAdjustMethod="BH")
delargMF <- enrichGO(l, OrgDb = org.Mm.eg.db, ont = "MF", universe = bg, pvalueCutoff=0.05, pAdjustMethod="BH")
delargCC <- enrichGO(l, OrgDb = org.Mm.eg.db, ont = "CC", universe = bg, pvalueCutoff=0.05, pAdjustMethod="BH")

BPdp <- dotplot(delargBP, showCategory=30) + ggtitle("dotplot Biological Process")
MFdp <- dotplot(delargMF, showCategory=30) + ggtitle("dotplot Molecular Function")
CCdp <- dotplot(delargCC, showCategory=30) + ggtitle("dotplot Cell Compartment")

pdf("dotplotBP.pdf", width=8, height = 12)
BPdp
dev.off()

pdf("dotplotMF.pdf", width=8, height = 12)
MFdp
dev.off()

pdf("dotplotCC.pdf", width=8, height = 12)
CCdp
dev.off()


#complete lists
BPdf <- as.data.frame(delargBP)
MFdf <- as.data.frame(delargMF)
CCdf <- as.data.frame(delargCC)

write_xlsx(BPdf, "BPdf.xlsx")
write_xlsx(MFdf, "MFdf.xlsx")
write_xlsx(CCdf, "CCdf.xlsx")


```


## revigo
process the output from the BP table at http://revigo.irb.hr/ to try to remove redundancies in the list
```{r andreasrevigo2}
#andreas now further curated the list and wants to use the top 25 
arev2 <- read_xlsx("Revigo_BP_OnScreenTable_nonrelevantred_sorted_ak.xlsx")

arev2r <- arev2[arev2$irrelevant=="no",]

arev2BPdf <- BPdf[BPdf$ID %in% arev2r$ID,]

#order
arev2BPdf <- arev2BPdf[order(arev2BPdf$p.adjust),]

arev2pdat <- arev2BPdf[1:25,]


# Split the ratio character strings into numerator and denominator
ratios_split <- strsplit(arev2pdat$GeneRatio, "/")

# Convert the numerator and denominator strings to numeric
numerator <- as.numeric(sapply(ratios_split, `[`, 1))
denominator <- as.numeric(sapply(ratios_split, `[`, 2))

# Calculate the ratios
ratios_calculated <- numerator / denominator

# Add the calculated ratios to the data frame as a new column
arev2pdat$calculated_ratios <- ratios_calculated

head(arev2pdat)


# export list with old pvalues
write_xlsx(arev2pdat, "Revigo_BP_OnScreenTable_nonrelevantred_sorted_ak_pvals.xlsx")




# make a dotplot for the top 25 by pvalue of this data
avrevplot <- ggplot(arev2pdat, aes(x=calculated_ratios, y=factor(Description, levels = arev2pdat[order(arev2pdat$calculated_ratios, decreasing = F),"Description"]), size=Count, colour= p.adjust)) + geom_point() +
  theme_paper() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        legend.position = "right") +
  scale_colour_gradient(high=farbe2, low=farbe1)
  

avrevplot

pdf("reviplot_akselect.pdf", width=4, height = 6)
avrevplot
dev.off()
```

## revigo unmodified
hier nochmal der unmodifizierte revigo output für die supplementary table (alte pvalues adden)
```{r revigounmod}
avrev_org <- read_xlsx("Revigo_BP_Table_full.xlsx")


arevorgBPdf <- BPdf[BPdf$ID %in% avrev_org$TermID,]

# export list with old pvalues
write_xlsx(arevorgBPdf, "Revigo_BP_Table_full_pval.xlsx")

```
